---
title: "Meeting 1"
author: "Rui Li"
date: "9/7/2020"
header-includes:
- \usepackage{amsgen,amsmath,amstext,amsbsy,amsopn,amssymb,mathabx,amsthm,bm,bbm}
- \usepackage[labelsep=space]{caption}
output:
  word_document: default
  html_document: default
subtitle: $\textbf{STOR 691H, FALL 2020}$
---
# Require packages
```{r message = False}
library(utils)
library(dplyr)
library(ggplot2)
library(maps)
library(stringr)
```

# Import the newest COVID-19 Dataset 
```{r}
#these libraries need to be installed
#library(utils)

#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
```

# Data cleaning and Summary
```{r}
#library("dplyr") require "dplyr" package for the use of %>%
covid.data = data %>%
              rename(Date="dateRep") %>%
              rename(Region="countriesAndTerritories") %>%
              mutate(Date=as.Date(Date,format="%d/%m/%y"))%>% 
              mutate(Region = str_replace_all(Region, "_", " "))

summary(covid.data)
head(covid.data)
```

```{r}
regions = unique(covid.data$Region)
region_w = unique(worldData$region)

missing_region = c()
for (i in 1:length(regions)) {
  if(!regions[i]%in%region_w){
    missing_region = c(regions[i],missing_region)
  }
}
missing_region = as.data.frame(missing_region); names(missing_region) = c("Missing Region")
world_region = as.data.frame(region_w); names(world_region) = c("World Region")
```

# Data Distribution of population
```{r}
#get a list of dates, countries, population, and continents
dates = unique(covid.data$Date)
regions = unique(covid.data$Region)
pops = unique(covid.data$popData2019)
continents = unique(covid.data$continentExp)

#population dataframe
pop.data = unique(data.frame(covid.data$Region,covid.data$popData2019,covid.data$continentExp))

names(pop.data)=c("Region","Population2019","Continents")

summary(pop.data)
head(pop.data)

#get the world polygon
worldData = map_data("world")
#plot world population based on color
ggplot() +
    geom_map(data = worldData, map = worldData,
                  aes(x = long, y = lat, group = group, map_id=region),
                  fill = "white", colour = "#7f7f7f", size=0.5) + 
    geom_map(data = pop.data, map=worldData,
                  aes(fill=Population2019, map_id=Region),
                  colour="#7f7f7f", size=0.5) +
    coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90)) +
    scale_fill_continuous(low="thistle2", high="darkred", guide="colorbar") +
    scale_y_continuous(breaks=c()) +
    scale_x_continuous(breaks=c()) +
    labs(fill="Population", title="Countries and Territories Population in 2019", x="", y="") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme_bw()
```
#mutate, str_replace_all() replace the undersocre. 
#Color Scale population in millions
#population density
#population/ square mile 2019

#https://bhaskarvk.github.io/user2017.geodataviz/notebooks/03-Interactive-Maps.nb.html\

#https://coronavirus.jhu.edu/map.html
#https://coronavirus.jhu.edu/data/mortality
#cases-death
#death/pop *100 - GDP economic - num.hospitals

#recreate and check; futher explore
#Combine three plots together
#Cases/pop incidence rate
#https://plotly.com/r/

#changing new cases for the time series analysis, setting model and do some predictions. 
#add limitations and compare different vaccine types, access the value of the vaccine. 

#ARIMA Models

# Data Distribution of cases
```{r}
#cases dataframe
case.data = data.frame(covid.data$Region,covid.data$cases) 
names(case.data) = c("Region","cases") 
case.data=aggregate(case.data$cases, by=list(Category=case.data$Region), FUN=sum)
names(case.data) = c("Region","cases")

summary(case.data)
head(case.data)

#get the world polygon
worldData = map_data("world")
#plot world population based on color
ggplot() +
    geom_map(data = worldData, map = worldData,
                  aes(x = long, y = lat, group = group, map_id=region),
                  fill = "white", colour = "#7f7f7f", size=0.5) + 
    geom_map(data = case.data, map=worldData,
                  aes(fill=cases, map_id=Region),
                  colour="#7f7f7f", size=0.5) +
    coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90)) +
    scale_fill_continuous(low="thistle2", high="darkred", guide="colorbar") +
    scale_y_continuous(breaks=c()) +
    scale_x_continuous(breaks=c()) +
    labs(fill="cases", title="Covid_19 Total Cases", x="", y="") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme_bw()
```

# Data Distribution of death
```{r}
#death dataframe
death.data = data.frame(covid.data$Region,covid.data$deaths) 
names(death.data) = c("Region","deaths") 
death.data=aggregate(death.data$deaths, by=list(Category=death.data$Region), FUN=sum)
names(death.data) = c("Region","deaths")

summary(death.data)
head(death.data)

#get the world polygon
worldData = map_data("world")
#plot world population based on color
ggplot() +
    geom_map(data = worldData, map = worldData,
                  aes(x = long, y = lat, group = group, map_id=region),
                  fill = "white", colour = "#7f7f7f", size=0.5) + 
    geom_map(data = death.data, map=worldData,
                  aes(fill=deaths, map_id=Region),
                  colour="#7f7f7f", size=0.5) +
    coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90)) +
    scale_fill_continuous(low="thistle2", high="darkred", guide="colorbar") +
    scale_y_continuous(breaks=c()) +
    scale_x_continuous(breaks=c()) +
    labs(fill="deaths", title="Covid_19 Total Deaths", x="", y="") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme_bw()
```

```{r}
#Other tries
centroids <- summarize(group_by(worldData, region),
                             x = mean(range(long)), y = mean(range(lat)))
names(centroids) <- c("Region","long","lat")

centroids <- centroids[centroids$Region%in%regions,]
head(centroids)

ggplot() + 
  geom_polygon(data = worldData,aes(x = long, y = lat, fill = region, group = group), color = "white") +  
  coord_fixed(1.3) +
  guides(fill=FALSE) + # do this to leave off the color legend
  geom_point(aes(x=long, y=lat), data = centroids, color = "red")
```

