---
title: "Senior Honor Thesis Project"
author: "Rui Li"
date: "11/28/2020"
header-includes:
- \usepackage{amsgen,amsmath,amstext,amsbsy,amsopn,amssymb,mathabx,amsthm,bm,bbm}
- \usepackage[labelsep=space]{caption}
output:
  html_document: default
subtitle: $\textbf{Coding & Modelings}$
---

**Require packages**
```{r message = FALSE, warning=FALSE, error=FALSE}
library(utils)
library(dplyr)
library(ggplot2)
library(maps)
library(stringr)
library(readr)
library(tidyverse)
library(readxl)
library(plotly)
library(MASS)
library(kableExtra)
library(broom)
library(stargazer)
```

**Cleaning Data**
```{r message = FALSE, warning=FALSE, error=FALSE}
#these libraries need to be installed
#library(utils)

#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")

#library("dplyr") require "dplyr" package for the use of %>%
covid.data = data %>%
  rename(Date="dateRep") %>%
  rename(Region="countriesAndTerritories") %>%
  mutate(Date=as.Date(Date,format="%d/%m/%y"))%>% 
  mutate(Region = str_replace_all(Region, "_", " "))%>% 
  group_by(Region) %>%
  mutate(cases_Diff = lag(cases)-cases) %>%
  mutate(deaths_Diff = lag(deaths)-deaths) %>% 
  relocate(cases_Diff, .after = cases) %>%
  relocate(deaths_Diff, .after = deaths) %>%
  drop_na(countryterritoryCode)
```

**Data Summary**
```{r message = FALSE, warning=FALSE, error=FALSE}
#population dataframe
pop.data = unique(data.frame(covid.data$Region,covid.data$popData2019,covid.data$continentExp,covid.data$countryterritoryCode))
names(pop.data)=c("Region","Population2019","Continents","Codes")
pop.data = na.omit(pop.data)
pop.data = unique(pop.data)

#cases dataframe
case.data = data.frame(covid.data$Region,covid.data$cases,covid.data$countryterritoryCode) 
names(case.data) = c("Region","cases","Codes") 
case.data=aggregate(case.data$cases, by=list(Category=case.data$Region), FUN=sum)
case.data = as.data.frame(cbind(case.data,unique(covid.data$countryterritoryCode)))
names(case.data) = c("Region","cases","Codes")
case.data = na.omit(case.data)

#death dataframe
death.data = data.frame(covid.data$Region,covid.data$deaths,covid.data$countryterritoryCode) 
names(death.data) = c("Region","deaths","Codes") 
death.data=aggregate(death.data$deaths, by=list(Category=death.data$Region), FUN=sum)
death.data = as.data.frame(cbind(death.data,unique(covid.data$countryterritoryCode)))
names(death.data) = c("Region","deaths","Codes")
death.data = na.omit(death.data)

#Death/Case dataframe
death_case.data = data.frame(case.data$Region,death.data$deaths/case.data$cases,case.data$Codes)
names(death_case.data) = c("Region","Deaths/Case Ratio","Codes")

#death/pop *100
death_pop.data = data.frame(pop.data$Region,death.data$deaths/pop.data$Population2019*100,pop.data$Codes)
names(death_pop.data) = c("Region","Values","Codes")

#Land Area
Land_Area <- read_excel("Land Area.xls")
names(Land_Area) = Land_Area[3,]
Land_Area = Land_Area[-(1:3),]
Land_Area = Land_Area[,c(1,2,62,63)]

#Complete 2018
for (i in 1:nrow(Land_Area)) {
  if(is.na(Land_Area$`2018`[i])){
    Land_Area$`2018`[i] = Land_Area$`2017`[i]
  }
}
Land_Area = na.omit(Land_Area)[,-3]
names(Land_Area)[3] = "Area"

combine = full_join(pop.data, Land_Area, by=c("Codes"="Country Code"))
combine = na.omit(combine)[,-5]
pop.square = na.omit(combine)
pop.square$Area = as.numeric(pop.square$Area)
```

**Ploting data**
```{r message=FALSE, warning=FALSE, error=FALSE}
#Population Ditribution
pop.fig <- plot_ly(pop.data, type='choropleth', locations=pop.data$Codes, z=log2(pop.data$Population2019), text=pop.data$Region, colorscale="Blues",reversescale =T)%>%
  layout(title = 'The Logarithm of World Population in 2019')%>%
  colorbar(title = "Population Rates",limits = c(15,31))
pop.fig

#Population Density Ditribution
pop.square.fig <- plot_ly(pop.square, type='choropleth', locations=pop.square$Codes, z=log(pop.square$Population2019/pop.square$Area), text=pop.square$Region, colorscale="Blues",reversescale =T)%>%
  layout(title = 'The Logarithm of World Population Density in 2019')%>%
  colorbar(title = "Density Rates",limits = c(-2,7))
pop.square.fig

#Case Ditribution
case.fig <- plot_ly(case.data, type='choropleth', locations=case.data$Codes, z=log2(case.data$cases), text=case.data$Region, colorscale="Reds",reversescale =F)%>%
  layout(title = 'The Logarithm of World Covid-19 Cases Number')%>%
  colorbar(title = "Cases Number",limits = c(4,24))
case.fig

#Death Ditribution
death.fig <- plot_ly(death.data, type='choropleth', locations=death.data$Codes, z=log2(death.data$deaths+1), text=case.data$Region, colorscale="Reds",reversescale =F)%>%
  layout(title = 'The Logarithm of World Covid-19 Deaths Number')%>%
  colorbar(title = "Deaths Number")
death.fig

#Death/Case Ditribution
death_case.fig <- plot_ly(death_case.data, type='choropleth', locations=death_case.data$Codes, z=death_case.data$`Deaths/Case Ratio`, text=death_case.data$Region, colorscale="Reds",reversescale =F)%>%
  layout(title = 'The Ratio of World Covid-19 Deaths to Cases Number')%>%
  colorbar(title = "Ratio Number",limits = c(0,0.1))
death_case.fig

#Death/pop*100 Ditribution
death_pop.fig <- plot_ly(death_pop.data, type='choropleth', locations=death_pop.data$Codes, z=death_pop.data$`Values`, text=death_pop.data$Region, colorscale="Reds",reversescale =F)%>%
  layout(title = 'The Ratio of World Covid-19 Death/Population*100')%>%
  colorbar(title = "Ratio Number")
death_pop.fig
```

**Modeling Preparation**
```{r message = FALSE, warning=FALSE, error=FALSE}
#Creat Density Variable
pop.density = pop.square %>%
  mutate(Density = Population2019/Area)
pop.density = pop.density[,c(4,6)]
#Creat data for time series
covid.time = covid.data[,c(1,6,8:9,11:12)]%>%
  rename(Codes="countryterritoryCode")%>%
  inner_join(pop.density)%>%
  mutate(Region = as.factor(Region))%>%
  mutate(group_id = as.integer(Region))
groups = length(unique(covid.time$Region))

#Lag cases and deaths Difference
for (k in 1:14) {
  CASE=NULL
  for (j in 1:groups) {
    CASE=c(CASE,lag(covid.time$cases_Diff[covid.time$group_id==j],k))
  }
  covid.time = cbind(covid.time,CASE)
}
names(covid.time)[9:22]=paste("CASE_DIFF",1:14,sep="-")

for (k in 1:14) {
  DEATH=NULL
  for (j in 1:groups) {
    DEATH=c(DEATH,lag(covid.time$deaths_Diff[covid.time$group_id==j],k))
  }
  covid.time = cbind(covid.time,DEATH)
}
names(covid.time)[23:36]=paste("DEATH_DIFF",1:14,sep="-")
```

**Modeling**

- Consider the 14 Lags of Cases Difference only.
```{r message = FALSE, warning=FALSE, error=FALSE}
#Cases on Cases
data1 = covid.time[,c(2,9:22)]
fit1.m1 <- lm(cases_Diff~., data=na.omit(data1))
fit2.m1 <- lm(cases_Diff~1, data=na.omit(data1))

step.m1 = stepAIC(fit2.m1,direction="both",scope=list(upper=fit1.m1,lower=fit2.m1),trace = FALSE)

summary(step.m1)
```

- Consider the 14 Lags of Cases and Deaths Difference only.
```{r message = FALSE, warning=FALSE, error=FALSE}
#Cases & Deaths on Cases
data2 = covid.time[,c(2,9:36)]
fit1.m2 <- lm(cases_Diff~., data=na.omit(data2))
fit2.m2 <- lm(cases_Diff~1, data=na.omit(data2))

step.m2 = stepAIC(fit2.m2,direction="both",scope=list(upper=fit1.m2,lower=fit2.m2),trace = FALSE)

summary(step.m2)

#Compare models
anova(step.m1,step.m2)
```

|Model Name|Res.DF|RSS|Df|Sum of Sq|F|Pr(>F)|
|----|----|----|----|----|----|----|
|1|52547|39.17|||||
|2|52534|38.78|13|388052910|40.438|<2.2e-16 ***|

- Consider the 14 Lags of Cases & Deaths Difference, and population.
```{r message = FALSE, warning=FALSE, error=FALSE}
#Cases, Deaths & Population on Cases
data3 = covid.time[,c(2,6,9:36)]
fit1.m3 <- lm(cases_Diff~., data=na.omit(data3))
fit2.m3 <- lm(cases_Diff~1, data=na.omit(data3))

step.m3 = stepAIC(fit2.m3,direction="both",scope=list(upper=fit1.m3,lower=fit2.m3),trace = FALSE)

summary(step.m3)

#Compare models
anova(step.m2,step.m3)
```

|Model Name|Res.DF|RSS|Df|Sum of Sq|F|Pr(>F)|
|----|----|----|----|----|----|----|
|2|52534|38.78|||||
|3|52533|38.77|1|11207507|15.187|9.75e-05 ***|

- Consider the 14 Lags of Cases & Deaths Difference, and density
```{r message = FALSE, warning=FALSE, error=FALSE}
#Cases, Deaths & Population on Cases
data4 = covid.time[,c(2,7,9:36)]
fit1.m4 <- lm(cases_Diff~., data=na.omit(data4))
fit2.m4 <- lm(cases_Diff~1, data=na.omit(data4))

step.m4 = stepAIC(fit2.m4,direction="both",scope=list(upper=fit1.m4,lower=fit2.m4),trace = FALSE)

summary(step.m4)

#Compare models
anova(step.m3,step.m4)
```

|Model Name|Res.DF|RSS|Df|Sum of Sq|F|Pr(>F)|
|----|----|----|----|----|----|----|
|3|52533|38.77|||||
|4|52534|38.78|-1|-11207507|15.187|9.75e-05 ***|

**Final Model**
```{r}
summary(step.m3)
```

**Model Presenting**
```{r message = FALSE, warning=FALSE, error=FALSE}
stargazer(step.m1, step.m2, title = "Results of Model 1 & 2", align = T,type = "text" )
stargazer(step.m3, step.m4, title = "Results of Model 3 & 4", align = T,type = "text" )
stargazer(step.m1, step.m2, step.m3, title = "Results of Model 1 & 2 & 3", align = T,type = "text")
```